Supplementary Information -- Morphological convergence as on-line lexical analogy

### 2020-03-03 supplementary_information/si.Rnw
final commit? yeah right

\noindent This is an Sweave file. The Rnw source file contains the code used to compile the pdf output file. It also contains additional code sections that help replicate the results. The files also contain pointers to further detail in the supplementary depository that accompanies the paper.

In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, the setup of (\ref{appendixgcm}) the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare. We illustrate the models using fits on our \emph {baseline} data.

@@

@@

The stimuli for the ESP experiment, 156 nonce verbs across four classes, are shown in Table 1 in the main text. Our formal criteria are based on the behavior of existing verbs in English and their categorization by \cite{bybee1982rules}, \cite{moder1992productivity}, and \cite{albright2003rules}. We made some adjustments to their categories. For instance, our DROVE class is a generalized version of Moder's class 4, described using the \textipa{[aI]}$\rightarrow{}$\textipa{[oU]} alternation but also including {\em weave}. (In contrast, Albright \& Hayes restrict this class to this alternation.) The verb classes could be defined in a number of ways and still be mostly consistent with the work of Moder, for instance. We have trialed a set of possible small changes and found that they have no major effect on our overall results.

@@

\begin{quote}
 \texttt{The implementation of the GCM is laid out in models/gcm/gcm.html}.
\end{quote}

@@

The irregular set consists of verb types in \textsc{Celex} that form their past tense according to the pattern captured by the class (such as an \{\textipa{[aI]},\textipa{[i]}\} $\rightarrow{}$ \textipa{[oU]} alternation). 
The regular set consists of verb types that have base forms that are similar to these irregular forms but have a regular {\em -ed} past tense form as well as miscellaneous regular verbs -- those that do not belong to our schemata . We narrow the regular set to monosyllabic forms. However, all polysyllabic irregular forms that could serve as a point of comparison are compounds based on monosyllabic forms (an example is {\em overwrite}, a compound form of irregular {\em write}). (A compound form might be more regular than a simplex form, but \textsc{Celex} will list both the regular and the irregular variant in both cases.) Table \ref{regexesforgcm} shows the descriptions of the verb classes using regular expressions.

@@

gcm_table = rbind(
  gcm_table,
  c('sum', 
    sum(gcm_table$`irregular set`),
    sum(gcm_table$`regular set`),
    unique(gcm_table$`miscellaneous regular verbs`)
    )
)

@@

Table \ref{gcmsetsize} shows the number of verbs in \textsc{Celex} that were used as training sets for our verb classes. The irregular set consists of forms that match the schema and are irregular. The regular set contains schema matches which are regular, in addition to miscellaneous regulars. The miscellaneous set consists of monosyllabic verbs that do not belong to any of the schemata and are regular. These are included in the regular training set of \textsc{ each} verb class.

@@

<<gcmtable, echo=F, eval=T, results=tex>>=
print(xtable::xtable(gcm_table, digits=2, include.rownames = FALSE, rowlabel="gcmsetsize", caption="Number of forms in each verb class, GCM training data"), include.rownames = F)

@@

\subsection{Output format}

@@

We standardize the regularity score to match the range of participant responses: [0,1]. The modified score is interpretable as the probability that the outcome will be regular in aggregated data. It is also appropriate to attribute this type of gradience to people's initial expectations about other people's behavior, on the assumption that people realistically encode the variability they have encountered.

@@

The Minimal Generalization Learner is implemented here from materials made available by Albright and Hayes \citep{albright2003rules}. These include their Segmental Similarity Calculator, implementing the natural class based similarity metric due to \cite{frisch1997similarity}, also used in the GCM implementation. Due to issues with the MGL code, we had to fit the MGL separately for participants.

\begin{quote}
 \texttt{The implementation of the MGL is laid out in models/mgl/mgl.html}.
\end{quote}

@@

Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other. As a consequence, the MGL rules apply to fewer forms than apparent from the table: 456 (instead of 617) in total.

@@

# mgl_count$form1 %>% unique %>% length

@@

@@

@@

Out of 156 test verbs in the ESP post-test, the \textsc{Celex}-trained MGL generates a {\em relevant regular rule} for every verb. It does not generate a {\em relevant irregular rule} for 28 verbs. These are all nonce verbs in the \textsc{KEPT} category (see\ Section 4.1 in the main text). In this category, irregular forms are derived from the stem through a vowel change (e.g. {\em greel} $\rightarrow{}$ {\em grelt}). The verbs missing the relevant irregular rule all have bases ending in $<$m$>$, $<$n$>$, or $<$l$>$. This is because, in our implementation, there is an insufficient number of verb types in the training set to support these irregular rules. Decreasing the cutoff criterion for the model leads to the generation of more of the currently `missing' irregular rules, but the overall model fit becomes worse (cf.\ below). Therefore, we keep the cutoff criterion and assume that the adjusted confidence of the irregular rule for these 28 verbs is zero. We then take the relative confidence of the regular rule as compared to the regular {\em and} the irregular rule for each verb and took this as the adjusted regular confidence of the given verb. (If the irregular rule is missing, the value of this adjusted regular confidence is 1.) This is given by the following equation:

@@

The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}. The {\em relevant regular rule} that generates the regular past tense is `$\emptyset{} \rightarrow{}$ \textipa{[d]} / \{\textipa{D}, \textipa{S}, \textipa{T}, \textipa{Z}, f, s, v, z\} \underline{\hspace{0.5cm}}'. The {\em structural description} indicates that this is a suffixation rule that can apply to forms that {\em end in an anterior fricative} (a natural class in our feature system).  The {\em raw confidence} of this rule is 0.98. This is because this rule applies to most forms in its scope (698/712). The {\em adjusted confidence} is very similar: 0.968. This is because this rule applies to a large number of forms overall. The {\em relevant irregular rule} that generates the irregular form is `\textipa{[aI]} $\rightarrow{}$ \textipa{[oU]} / \{\textipa{D}, \textipa{Z}, \textipa{dZ}, d, l, n, r, z\} \underline{\hspace{0.5cm}}v'. It applies to \textipa{[aI]} in the nucleus {\em preceded by voiced anterior consonant and followed by \textipa{[v]}}. In \textsc{Celex},the rule applies to three forms ({\em drive, strive, dive}) and fails to apply to five ({\em arrive, thrive, contrive, rive, connive}). Its {\em raw confidence} is 0.375. Its {\em adjusted confidence} is slightly lower (0.366). This is because it applies to a smaller number of forms overall. The {\em relative (adjusted) confidence} of the predicted regularity of {\em splive} is $0.98/(0.98+0.37)=0.73$. (The respective rescaled value is 0.363 -- which, in itself, is not very informative.)

@@

As we see, however, the MGL also focuses the training set. The structural descriptions of the forms only cover a fraction of all verbs in the starting dictionary, organized by shared groups of segments. While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence} . In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class. The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mgltable}.

@@

# glue on column sums
mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))

@@

@@

Our regression model selection in this article proceeded as follows. Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}. Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function. However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.

We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors. We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.

In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}\footnote{In the SI, all p values rounded to 0.00 should be taken to indicate p $<$ 0.001.}.  Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}. We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model. A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions. However, we elect to take a more conservative approach by not including a potentially spurious main effect. Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case. The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.9.

@@

We fit the \textsc{Celex-GCM} and the \textsc{Celex-MGL} on the baseline data to see how well they predict regularization responses by the baseline participants. The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, if slightly. Both models contribute independently to explaining variation in the data.

@@

The concordance indices of the \textsc{Celex}-MGL (\Sexpr{C_CELEX_mgl_features}) and the \textsc{Celex}-MGL (\Sexpr{C_CELEX_gcm_features}) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}. Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.

@@

We fit the \textsc{Celex-GCM},the \textsc{Celex-MGL}, as well as the Individual-GCM and the Individual-MGL on the post-test data. Recall that the \textsc{Celex}-based models were trained on verbs in \textsc{Celex}. The individual-based models were trained for each participant, on \textsc{Celex} verbs as well as the co-player's responses in the ESP test. They were tested on the post-test for each participant, individually.

@@

<<indiv_diff2, echo=F,eval=T, results=tex>>=
load('models/si_indiv_m6.rda')
indiv_m6_sum = niceCoefTable(indiv_m6)
print(xtable::xtable(indiv_m6_sum, digits = 3, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'), include.rownames = F)
@

@@

@@

Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ \Sexpr{chisq1}) or individual-GCM ($\chi^2 = $ \Sexpr{chisq2}) as predictors. It is not significantly better if we include individual-MGL ($\chi^2 = $ \Sexpr{chisq3}) or {\sc Celex}-GCM ($\chi^2 = $ \Sexpr{chisq4}).

@@

# plot this, x is baseline, y is posttest.
# colour AND shape are lexical typicality.
# split panels across raw data / gcm / mgl.
# add loess smooths.
p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = co_player_lexical_typicality, shape = co_player_lexical_typicality, linetype = co_player_lexical_typicality)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  # scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline regularization') +
  ylab('verb ESP regularization') +
  guides(linetype = F) +
  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")

@@

\begin{figure}
<<model_figure2, fig=T,echo=F,eval=T, width = 12.5, height = 4, fig.margin=T, fig.align = ‘center’>>=
p1
@
\caption{Baseline and individual regularization weights: participants, GCM, MGL}
\label{bigfigure}
\end{figure}

@@

While the loess smooths are likely slightly overfit, they show the general trend in the raw data, which is echoed in the regression analysis in the main text: the `typical' trend of baseline-post-test difference is stronger than either the "random" or "reversed" trends. These are not very much different from each other. The differences are gradual. The GCM, while generally overestimating regularity, captures this gradual shift, since verb behavior is shifted individually in the post-test. The MGL also overestimates regularization. Since rules affect groups of verbs, rather than individual verbs, the post-test shifts are abrupt. The "typical" trend continues climbing after a longer plateau. However, the `random' and `reversed' trends trail off, where the "reversed" trend turns negative towards the higher tail of the distribution.

>>>>

\noindent In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, the setup of (\ref{appendixgcm}) the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare. We illustrate the models using fits on our \emph {baseline} data.

@@

\begin{quote}
This is an Sweave file. The Rnw source file contains the code used to compile the pdf output file. It also contains additional code sections that help replicate the results. The file also contains pointers to further detail in the supplementary repository that accompanies the paper. 
\end{quote}

@@

The stimuli for the ESP experiment, 156 nonce verbs across four classes, are shown in Table 1 in the main text. Our formal criteria are based on the behavior of existing verbs in English and their categorization by \cite{bybee1982rules}, \cite{moder1992productivity}, and \cite{albright2003rules}. We made some adjustments to their categories. For instance, our DROVE class is a generalized version of Moder's class 4, described using the \textipa{[aI]}$\rightarrow{}$\textipa{[oU]} alternation but also including {\em weave}. (In contrast, Albright \& Hayes restrict this class to the \textipa{[aI]}$\rightarrow{}$\textipa{[oU]} alternation.) The verb classes could be defined in a number of ways and still be mostly consistent with the work of Moder, for instance. We have trialed a set of small changes in defining the verb categories and found that they have no major effect on our overall results.

@@

\begin{quote}
 The implementation of the GCM is laid out in models/gcm/gcm.html.
\end{quote}

@@

The irregular set consists of verb types in \textsc{Celex} that form their past tense according to the pattern captured by the class (such as an \{\textipa{[aI]},\textipa{[i]}\} $\rightarrow{}$ \textipa{[oU]} alternation). 
The regular set consists of verb types that have base forms that are similar to these irregular forms but have a regular {\em -ed} past tense form as well as miscellaneous regular verbs -- those that do not belong to any of our schemata. We narrow the regular set to monosyllabic forms. However, all polysyllabic irregular forms that could serve as a point of comparison are compounds based on monosyllabic forms (an example is {\em overwrite}, a compound form of irregular {\em write}. A compound form might be more regular than a simplex form, but \textsc{Celex} will list both the regular and the irregular variant in both cases.) Table \ref{regexesforgcm} shows the descriptions of the verb classes using regular expressions.

@@

gcm_table = rbind(
  gcm_table,
  c('count of unique forms', 
    sum(gcm_table$`irregular set`),
    sum(gcm_table$`regular set`),
    unique(gcm_table$`miscellaneous regular verbs`)
    )
)

@@

Table \ref{gcmsetsize} shows the number of verbs in \textsc{Celex} that were used as training sets for our verb classes. The irregular set consists of forms that match the schema and are irregular. The regular set contains schema matches which are regular, in addition to miscellaneous regulars. The miscellaneous set consists of monosyllabic verbs that do not belong to any of the schemata and are regular. These are included in the regular training set of each verb class.

@@

<<gcmtable, echo=F, eval=T, results=tex>>=
print(xtable::xtable(gcm_table, digits=2, include.rownames = FALSE, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"), include.rownames = F)

@@

\subsection{Output format \label{gcm_scale}}

@@

We rescale the regularity score to match the range of participant responses: [0,1]. The modified score is interpretable as the probability that the outcome will be regular in aggregated data. It is also appropriate to attribute this type of gradience to people's initial expectations about other people's behavior, on the assumption that people realistically encode the variability they have encountered.

@@

The Minimal Generalization Learner is implemented here from materials made available by Albright and Hayes \citep{albright2003rules}. These include their Segmental Similarity Calculator, implementing the natural class based similarity metric due to \cite{frisch1997similarity}, also used in the GCM implementation. Due to issues with the MGL code, we had to fit the MGL manually, using the graphical interface, for each separate participant. The pyautogui library in python was used to automate this, the code is available in the supplementary repository.

\begin{quote}
The implementation of the MGL is laid out in models/mgl/mgl.html.
\end{quote}

@@

@@

@@

# we also want the TOTAL number of related forms and exceptions across rules.

mgl_count_total1 = mgl_count %>% 
  dplyr::select(related_list) %>% 
  tidyr::unnest(related_list) %>% # we now have the related forms as one verb per line, which is tidy. but the verbs repeat, since the same verb can be a related form for several target verbs
  dplyr::distinct() %>% # so we distinct them!
  dplyr::mutate(
    type = 'related forms' # by the way, this is related forms
  ) %>% 
  dplyr::rename(
   forms = related_list 
  )

mgl_count_total2 = mgl_count %>% 
  dplyr::select(exception_list) %>% 
  tidyr::unnest(exception_list) %>% # we now have the related forms as one verb per line, which is tidy. but the verbs repeat, since the same verb can be a related form for several target verbs
  dplyr::distinct() %>% # so we distinct them!
  dplyr::mutate(
    type = 'exceptions' # by the way, this is related forms
  ) %>% 
  dplyr::rename(
   forms = exception_list 
  )

all_forms_n_mgl = mgl_count_total2 %>% 
  dplyr::bind_rows(mgl_count_total1) %>% 
  dplyr::distinct(forms) %>% 
  dplyr::summarise(n =n()) %>% 
  dplyr::pull(n)

@@

Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other. As a consequence, the MGL rules apply to fewer forms than apparent from the table: \Sexpr{all_forms_n_mgl} in total.

@@

Out of 156 test verbs in the ESP post-test, the \textsc{Celex}-trained MGL generates a {\em relevant regular rule} for every verb. It does not generate a {\em relevant irregular rule} for 28 verbs. These are all nonce verbs in the \textsc{KEPT} category (see\ Section 4.1 in the main text). In this category, irregular forms are derived from the stem through a vowel change (e.g. {\em greel} $\rightarrow{}$ {\em grelt}). The verbs missing the relevant irregular rule all have bases ending in $<$m$>$, $<$n$>$, or $<$l$>$. This is because, in our implementation, there is an insufficient number of verb types in the training set to support these irregular rules. Decreasing the cutoff criterion for the model leads to the generation of more of the currently `missing' irregular rules, but the overall model fit becomes worse (cf.\ below). Therefore, we keep the cutoff criterion and assume that the adjusted confidence of the irregular rule for these 28 verbs is zero. We then take the relative confidence of the regular rule as compared to the regular {\em and} the irregular rule for each verb and take this as the adjusted regular confidence of the given verb. (If the irregular rule is missing, the value of this adjusted regular confidence is 1.) This is given by the following equation:

@@

The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}. The {\em relevant regular rule} that generates the regular past tense is `$\emptyset{} \rightarrow{}$ \textipa{[d]} / \{\textipa{D}, \textipa{S}, \textipa{T}, \textipa{Z}, f, s, v, z\} \underline{\hspace{0.5cm}}'. The {\em structural description} indicates that this is a suffixation rule that can apply to forms that {\em end in an anterior fricative} (a natural class in our feature system).  The {\em raw confidence} of this rule is 0.98. This is because this rule applies to most forms in its scope (698/712). The {\em adjusted confidence} is very similar: 0.968. This is because this rule applies to a large number of forms overall. The {\em relevant irregular rule} that generates the irregular form is `\textipa{[aI]} $\rightarrow{}$ \textipa{[oU]} / \{\textipa{D}, \textipa{Z}, \textipa{dZ}, d, l, n, r, z\} \underline{\hspace{0.5cm}}v'. It applies to \textipa{[aI]} in the nucleus {\em preceded by voiced anterior consonant and followed by \textipa{[v]}}. In \textsc{Celex},the rule applies to three forms ({\em drive, strive, dive}) and fails to apply to five ({\em arrive, thrive, contrive, rive, connive}). Its {\em raw confidence} is 0.375. Its {\em adjusted confidence} is slightly lower (0.366). This is because it applies to a smaller number of forms overall. The {\em relative (adjusted) confidence} of the predicted regularity of {\em splive} is $0.98/(0.98+0.37)=0.73$. (As with the GCM predictions -- see Section \ref{gcm_scale} of this SI -- we rescale the MGL predictions as well for statistical analysis. The respective rescaled value is 0.363.)

@@

@@

# we get a total rule count as well. since rules overlap, this isn't a sum of the rule counts above

total_rules = mgl_baseline_rules_raw %>% # take raw rules
  dplyr::filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
    !is.na(category),
    !is.na(rule.type)
  ) %>% 
  dplyr::mutate( # make a rule column of the form A -> B / C
    rule = paste0(A, ' ->', B, ' / ', Pfeat)
  ) %>% 
  dplyr::distinct(rule,rule.type) %>% # keep DISTINCT rules
  dplyr::count(rule.type)

irreg_total_count = total_rules[total_rules$rule.type == 'irregular',]$n
reg_total_count = total_rules[total_rules$rule.type == 'regular',]$n

As we see, however, the MGL also focuses the training set. The structural descriptions of the forms only cover a fraction of all verbs in the starting dictionary, organized by shared groups of segments. While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence}. In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class. The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mglruletable}. Note that the regular rules themselves overlap: A regular rule that accounts for \emph{weep-weeped} can also account for \emph{burn-burned}. In total, there are \Sexpr{irreg_total_count} irregular rules and \Sexpr{reg_total_count} regular rules posited by the MGL on the {\sc Celex} verbs.

<<mglruletable, echo=F, eval=T, results=tex>>=
print(xtable::xtable(mgl_rule_count, digits=2, label="mglruletable", caption="Number of regular and irregular rules in each verb class, posited by the MGL for the training data (Note that the regular rules themselves overlap)"), include.rownames = F)

@@

# here are the effect estimate figures from the paper:
# plot(effects::effect('co_player_regularisation_rate', lme.fit1), xlab = 'co-player regularization shift', ylab = 'player post-test regularization', main = 'Effect of co-player rate of regularization on pos-test responses')

# plot(effects::effect('verb_baseline_average:co_player_lexical_typicality', lme.fit1), multiline = T, lines=list(lty=c(1,2,3)), xlab = 'verb baseline average', ylab = 'player post-test regularization', main = 'Effect of co-player lexical typicality on pos-test responses')

@@

Our regression model selection in this article proceeded as follows. Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}. Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function. However, our model tables (e.g., Tables 5 and 6 in the main text) include the Wald significance levels generated by lme4 summaries (* p$<$0.05; ** p$<$0.01; *** p$<$0.001).  
We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors. We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.

In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}.  Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}. We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model. A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions. However, we elect to take a more conservative approach by not including a potentially spurious main effect. Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case. The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.9.

@@

We fit the \textsc{Celex-GCM} and the \textsc{Celex-MGL} on the baseline data to see how well they predict regularization responses by the baseline participants. The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, albeit only slightly. Both models contribute independently to explaining variation in the data.

@@

The concordance indices of the \textsc{Celex}-MGL (\Sexpr{C_CELEX_mgl_features}) and the \textsc{Celex}-MGL (\Sexpr{C_CELEX_gcm_features}) on the baseline data are similar, with some edge retained by the MGL, as in the case of \cite{albright2003rules}. Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.

@@

We fit the \textsc{Celex-GCM}, the \textsc{Celex-MGL}, as well as the Individual-GCM and the Individual-MGL on the post-test data. Recall that the \textsc{Celex}-based models were trained on verbs in \textsc{Celex}. The individual-based models were trained for each participant, on \textsc{Celex} verbs as well as the co-player's responses in the ESP test. They were tested on the post-test for each participant, individually.

@@

<<indiv_diff2, echo=F,eval=T, results=tex>>=
load('models/si_indiv_m6.rda')
indiv_m6_sum = niceCoefTable(indiv_m6)
print(xtable::xtable(indiv_m6_sum, digits = 2, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'), include.rownames = F)
@

@@

p3 = anova(comp_fit1, comp_fit1d)[2,8] %>% round(2) # with individual mgl
p4 = anova(comp_fit1, comp_fit1e)[2,8] %>% round(2) # with celex gcm

@@

Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ \Sexpr{chisq1} ***) or individual-GCM ($\chi^2 = $ \Sexpr{chisq2} ***) as predictors. It is not significantly better if we include individual-MGL ($\chi^2 = $ \Sexpr{chisq3}, p = \Sexpr{p3}) or {\sc Celex}-GCM ($\chi^2 = $ \Sexpr{chisq4}, p = \Sexpr{p4}).

@@

# plot this, x is baseline, y is posttest.
# colour AND shape are lexical typicality.
# split panels across raw data / gcm / mgl.
# add loess smooths.
p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = co_player_lexical_typicality, shape = co_player_lexical_typicality, linetype = co_player_lexical_typicality)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type, ncol = 2) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  # scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline regularization') +
  ylab('verb ESP regularization') +
  guides(linetype = F) +
  theme_bw() +
  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")

ggsave('si_figure.pdf', plot = p1, width = 9, height = 7)

@@

\begin{figure}
\centering
\includegraphics[width = \textwidth]{si_figure.pdf}
\caption{Baseline and individual regularization weights: participants, GCM, MGL}
\label{bigfigure}
\end{figure}

@@

While the loess smooths are likely slightly overfit, they show the general trend in the raw data, which is echoed in the regression analysis in the main text: the `typical' trend of baseline-post-test difference is stronger than either the `random' or `reversed' trends. These are not very much different from each other. The differences are gradual. The GCM, while generally overestimating regularity, captures this gradual shift, since verb behavior is shifted individually in the post-test. The MGL also overestimates regularization. Since rules affect groups of verbs, rather than individual verbs, the post-test shifts are abrupt. The `typical' trend continues climbing after a longer plateau. However, the `random' and `reversed' trends trail off, where the `reversed' trend turns negative towards the higher tail of the distribution.
